---
author: "Hector Roux de Bézieux"
date: '`r format(Sys.time(), "%d %B , %Y")`'
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
    code_download: TRUE
    
params:
  dataset: baron 
  title: "Analysis of the Baron dataset"
---
---
title: `r params$title`
---

```{r load packages, include=F}
library(knitr)
opts_chunk$set(
  fig.pos = "!h", out.extra = "", warning = F, message = F,
  fig.width = 5, fig.align = "center", echo = F
)
libs <- c("here", "dplyr", "ggplot2", "tidyr", "stringr", "readr", "cowplot",
          "clusterExperiment", "mclust", "RColorBrewer", "purrr", "Dune",
          "png")
suppressMessages(
  suppressWarnings(sapply(libs, require, character.only = TRUE))
)
rm(libs)
Dune <- lapply(1:3, function(i){
  read_csv(here("Data", "Dune", paste0(params$dataset, "_comp", i, "_Dune.csv")))
})
```

# EDA

We can first visualize the data after dimension reduction with zinbWave. We then use t-SNE and color the cells with the allen labels. Since we use various values for the k parameter, we plot two representations with the most extreme values of k.

```{r t-SNE plots, fig.width=18, fig.height=9}
plots <- list.files(here("Figures", "EDA"))
plots <- plots[str_detect(plots, params$dataset)]
plots <- plots[str_detect(plots, "png")]
plot.new()
plot.window(0:1, 0:1)
rasterImage(readPNG(here("Figures", "EDA", plots[1])), 0, 0, .5, 1)
rasterImage(readPNG(here("Figures", "EDA", plots[length(plots)])), 0.5, 0, 1, 1)
```


# Reduction in the number of clusters

```{r Imp no allen}
prePost <- function(df) {
  n_clust <- map_df(df, n_distinct) %>%
    select(-cells) %>%
    pivot_longer(everything(), names_to = "method", values_to = "n_clus") %>%
    dplyr::mutate(step = word(method, 2, sep = "-"),
                  method = word(method, 1, sep = "-")) %>%
    filter(step %in% c("Initial", "Final")) %>%
    dplyr::mutate(step = factor(step, levels = c("Initial", "Final")))
  p <- ggplot(n_clust, aes(x = method, y = n_clus, fill = step)) +
    geom_bar(stat = "identity", position = "dodge") +
    theme_classic() +
    scale_fill_viridis_d(option = "E") +
    labs(x = "Clustering Methods", y = "Number of clusters", fill = "") +
    ggtitle("Reduction in number of clusters with ARI merging")
  return(p)
}
prePost(Dune[[1]])
```

# Improvement in ARI

```{r ARI imp no allen cell}
ARI <- function(df) {
  keep <- str_detect(colnames(df), "Initial") |
          str_detect(colnames(df), "Final")
  df <- df[, keep]
  plotARIs(df[, 1:3])
  plot_grid(plotARIs(df[, 1:3]),
            plotARIs(df[, 4:6]))
}
ARI(Dune[[1]])
```

# Comparison with the reference

## After ARI merging

```{r comp cell}
ref_clusters <- read.csv(here("data", params$dataset,
                              paste0(params$dataset, "_meta.csv")))

final <- map_df(mergers$currentMat, adjustedRandIndex, y = allen_clusters$subclass_label) %>%
  pivot_longer(everything(), names_to = "clus_method", values_to = "ARI")
initial <- map_df(mergers$initialMat, adjustedRandIndex, y = allen_clusters$subclass_label) %>%
  pivot_longer(everything(), names_to = "clus_method", values_to = "ARI")
  
Monocle <- read.csv(here("data", "SingleMethod", paste0(params$dataset, "_Monocle.csv"))) %>%
  select(-X)
colnames(Monocle) <- paste0("Monocle_", str_remove(colnames(Monocle), "^k_"))
colnames(Monocle)[ncol(Monocle)] <- "cells"
sc3 <- read.csv(here("data", "SingleMethod", paste0(params$dataset, "_SC3.csv"))) %>%
  select(-X)
colnames(sc3) <- paste0("sc3_",
                        str_remove(colnames(sc3), "X") %>% str_replace("\\.", "-"))
colnames(sc3)[ncol(sc3)] <- "cells"
Seurat <- read.csv(here("data", "SingleMethod", paste0(params$dataset, "_Seurat.csv"))) %>%
  select(-X)
colnames(Seurat) <- paste0("Seurat_", str_remove(colnames(Seurat), "^X"))
colnames(Seurat)[ncol(Seurat)] <- "cells"
  
singles <- full_join(sc3, Seurat) %>% full_join(Monocle) %>%
  arrange(cells) %>% select(-cells) %>% as.list() %>%
  map_df(., adjustedRandIndex, y = allen_clusters$subclass_label) %>%
  pivot_longer(everything(), names_to = "method", values_to = "ARI") %>%
  dplyr::mutate(clus_method = word(method, 1, sep = "_"),
                level = word(method, 2, sep = "_"))
ggplot(singles, aes(x = clus_method, y = ARI)) +
  geom_violin() +
  theme_classic() +
  geom_point(
    data = bind_rows("Final" = final, "Initial" = initial, .id = "step"),
    aes(col = step)) +
  scale_color_brewer(type = "qual")
```


## After a last consensus step

```{r final consensus}
clusters <- mergers$currentMat
r <- which(colnames(clusters) == "RsecT")
if (all.equal(integer(0) ,r) != TRUE) {
  clusters[,"Rsec"] <- assignRsec(mergers) 
}
clusters <- as.matrix(clusters) 

cellsConsensus <- Consensus(clusMat = clusters,
                            large = (type != "Smart-Seq"))
consensus <- cbind(cellsConsensus,
                   allen_clusters$subclass_label,
                   clusters)

colnames(consensus)[c(1, 2)] <- c("Consensus", "Allen\nsubclass")
plotClusters(object = consensus)
title("Consensus Merging with proportion\nand comparison with the allen subclass.")

rm(clusters, cellsConsensus, consensus)
```

## Stopping during the merging

```{r consensus with intermediate}
midMat <- intermediateMat(merger = mergers,
                          p = .9)
r <- which(colnames(midMat) == "RsecT")
if (all.equal(integer(0) ,r) != TRUE) {
  midMat[,"Rsec"] <- assignRsec(mergers, p = .9)
}
midMat <- as.matrix(midMat)

cellsConsensus <- Consensus(clusMat = midMat,
                            large = (type != "Smart-Seq"))
consensus <- cbind(cellsConsensus,
                   allen_clusters$subclass_label,
                   midMat) %>% as.data.frame()

colnames(consensus)[c(1, 2)] <- c("Consensus", "Allen\nsubclass")
plotClusters(object = consensus %>% as.matrix())
title("Consensus Merging when stopping early")

rm(cellsConsensus, consensus, midMat)
```

# Seurat ARI Param

```{r seurat param}
plot.new()
plot.window(0:1, 0:1)
rasterImage(readPNG(
  here("Figures", type, paste0(params$dataset, "_seurat_ARI.png"))),
  0, 0, 1, 1)
```
